\boldmath Primordial Black Hole Formation in a Dust Bouncing Model

### 2024-12-03 main.tex
Breaking lines at col 88. Fixing minor typos.

\section{Linear Scalar Perturbations}
\label{linearpert}
While the universe appears homogeneous and isotropic on large scales, the FLRW metric falls short of providing a precise description of our universe, which presents inhomogeneities that are associated with structure formation, e.g. galaxy clusters, black holes, stars, and others. To characterize the physical universe, we study linear scalar perturbations around our flat contracting background metric described in the previous section. We will then investigate how said perturbations may seed the formation of primordial black holes in the subsequent section. We will from now on characterize background quantities with an overbar and perturbed (physical) quantities without it.

@@

Note from Eq.~\eqref{deltarhoinvariant} that if we choose a gauge where $\mathcal{V} =0$, $\Tilde{\delta\rho}$ becomes the physical density perturbation, which leads to an easier interpretation of this quantity. Also, the choice of $\mathcal{V} =0$ will be well suited to connect our perturbation theory with the Lemaitre-Toman-Bondi metric discussed in App.~\ref{appgauge}.  Other works have defined the gauge invariant density perturbation as
\begin{align}
	\label{newtoniandeltarho}
	\Tilde{\delta\rho}^{N} & \equiv \delta\rho - \delta\sigma \dot{\Bar{\rho}}
	,\end{align}
so that this quantity becomes the physical variable in the Newtonian Gauge, where $\delta\sigma = 0$ and $\Phi = \phi$. However, in Ref.~\cite{vitenti2012large}, it was shown that for bounce models with long contracting phases, regardless of the bounce type, the Bardeen perturbation $\Phi$ grows larger than one and invalidates the perturbative series, as $\phi$ also grows larger than one in the Newtonian Gauge. Additionally, the definition in Eq.~\eqref{newtoniandeltarho} leads to
\begin{align}
	\zeta & = \Phi + \frac{2\bar{D}^2 \Phi}{3  \kappa (\bar{\rho}+\bar{p})}+  \frac{{\tilde{\delta\rho}^N}} {3(\bar{\rho}+\bar{p})}
	,\end{align}
which implies that $\tilde{\delta\rho}^N$ grows with $\Phi$ and thus has larger spectra
as well as in this gauge. Hence, the Newtonian gauge is not a valid choice for bounce
models as it would lead to miss-calculation of the physical quantities whose values
would be inflated. We shall avoid this choice in this work and stick with the definition
in Eq.~\eqref{deltarhoinvariant}. We now compute the density contrast modes, which
require the perturbative equations of motion of the model.

@@

\subsection{Quantization}
%%%%%%%%%%%%%%%%%%%%%%%%

@@

A widely used prescription to define such a vacuum state is known as the adiabatic
vacuum prescription~\cite{birrell1984quantum, mukhanov2007introduction}. Its main idea is
to set initial conditions by demanding that the mode functions $\zeta_{k}(t)$ coincide
with their ${\cal N}$ order adiabatic approximation, $^{({\cal N})}\zeta_{k}(t)$. This
is implemented if one chooses the initial conditions~\cite{birrell1984quantum}
\begin{align}
	\label{v_init}
	{ }^{(\mathcal{N})}	\zeta_k(t_0) & = \frac{1}{\sqrt{m(t_0)\omega_k}(t_0)}e^{i\alpha_k(t_0)}~\text{and}~ { }^{(\mathcal{N})}	\dot{\zeta}_k(t_0) = im(t_0)\omega_k(t_0)\zeta_k(t_0)
	,\end{align}
where $\omega_k^2(t) = \frac{c_s^2 k^2}{\bar{a}^2}$, $m(t) = 2z^2$ and $\alpha_k$ is defined by the relation
\begin{align}
	{ }^{(\mathcal{N})}	\zeta_k(t_0) \dot{\alpha}_k(t_0) & = 1.
\end{align}

@@

As known from the literature, the dust bouncing model has an approximate scale-invariant spectrum, that is
\begin{align}
	\label{powerlaw}
	P_{{\zeta}}(k) & \approx A k^{n_s-1}
	,\end{align}
where $A$ is a constant and $n_s$ is the spectral index. This is a general feature of single barotropic fluid quantum bouncing models, and the spectral index is related to the equation of state parameter $w$ by\footnote{For an explicit semi-analytic derivation, see \cite{nelson_peter_bouncing_original}.}
%
\begin{equation}
	n_{s}(w) = 1 + \frac{ 12w }{ 1 + 3w }\, ,
\end{equation}
%
which is nearly scale invariant for $|w| \ll 1 $. However, the usual positive values of $w$ lead to a blue tilted spectrum $n_{s} > 1$, which differ from CMB observations \cite{planck_inflation_constraints}, consistent with the inflationary scenario prediction of a red tilted spectrum with $n_{s} \approx 0.96$. Although the initial power spectrum is not consistent with observations, one must recall that our model is considering a pure dark matter-dominated universe, neglecting the effects of radiation. In particular, it has been suggested \cite{nelson2021bouncing} that the inclusion of radiation may lead to a red almost scale-invariant spectrum. Therefore, this model must be understood as a first approximation to a more complete model, which is still under development. Furthermore, since most of the modes that influence our universe have crossed the Hubble horizon during dust domination, this means that this model, even with its simplicity, may still cover relevant information for the future complete model.

This result should not advocate for the prediction failure of the contracting scenario
since some works also suggest a blue-tilted spectrum for inflationary
models~\cite{Wang2014, Cai2015}. For instance, in Ref.~\cite{Kuroyanagi2021} they
discuss the possibility of explaining the recent NANOGrav results with a blue-tilted
specrta~\cite{Wu2023}. Also, the spectra are only dictated by a power law for narrow
intervals of momenta, which implies that corrections to Eq.~\eqref{powerlaw} must be
applied. Nonetheless, we see that the contracting phase produces a nearly
scale-invariant spectrum. In the next section, we begin to analyze the primordial black
hole formation seeded by the scalar perturbations in our cosmological background.

@@

It is indeed expected that a fluid with no pressure leads to a total collapse since no forces are opposing the collapse and dust only interacts via gravity. Thus all dark matter that we see today would indeed be contained in primordial black holes. However, this hypothesis was already refuted by some works by constraining the PBH abundance in dark matter based on observational effects (see Refs.~\cite{Villanueva2021, Carr2021}). Hence we must consider a dust fluid with a non-vanishing equation of state.
\subsection{Filtered Variance}
Let us now compute the variance which will determine the distribution for PBH. Since sub-Jeans and super-Hubble scales do not contribute to the PBH formation that we are interested in, we want to compute the integral in  Eq.~\eqref{eqsigmar} between $k_H$ and $k_j$. To do so, we first need to evaluate $P_{\tilde{\delta}}(k)$ using Eqs.~\eqref{deltarho}-\eqref{phiz} and the field expansions in Eqs.~\eqref{vexp} and~\eqref{pvexp}. Explicitly,
\begin{align}
	\label{psiq}
	\hat{\Psi}(\eta,\textbf{x}) & =\int\frac{\mathrm{d}^3\textbf{k}}{(2\pi)^{\frac{3}{2}}}
	\frac{\bar{H}}{2\Bar{a}k^2}\left[\Pi_{\zeta_{k}} e^{i\textbf{k}\textbf{x}} a_{k} +
	\Pi_{\zeta_{k}}^{ *}e^{-i\textbf{k}\textbf{x}} a_{k}^{\dagger}\right]
	,\end{align}
\begin{align}
	\bar{D}^2\hat{\Psi}(\eta,\textbf{x}) & =
	-\int\frac{\mathrm{d}^3\textbf{k}}{(2\pi)^{\frac{3}{2}}}
	\frac{\bar{H}}{2a^3 }\left[\Pi_{\zeta_{k}} e^{i\textbf{k}\textbf{x}} a_{k} +
	\Pi_{\zeta_{k}}^{*}e^{-i\textbf{k}\textbf{x}} a_{k}^{\dagger}\right]
\end{align}
and
\begin{align}
	\hat{\tilde{\delta\rho}}(\textbf{x}) & =
	\int_{-\infty}^{\infty}\frac{\mathrm{~d}^3\textbf{k}}{(2\pi)^{\frac{3}{2}}}
	\left\{ \left(\frac{\bar{H}}{\kappa \bar{a}^3}\right)\Pi_{\zeta_{k}} e^{i\textbf{k}\textbf{x}} a_{k} +\right. \nonumber                                     \\
	                                     & \left. \left(\frac{\bar{H}}{\kappa a^3}\right)\Pi_{\zeta_{k}}^{*} e^{-i\textbf{k}\textbf{x}} a_{k}^{\dagger}\right\}
	.\end{align}
Thus, the two-point function and the variance are given by
\begin{align}
	\label{2point}
	\left<\hat{\Tilde{\delta}}(\textbf{x})\hat{\Tilde{\delta}}(\textbf{y})\right> & =
	\frac{1}{2\pi^2} \int_0^\infty \frac{\mathrm{~d}k}{k}~\left[P_{\tilde{\delta}}(k)
		\frac{\sin{kR}}{kR} \right]
\end{align}
and
\begin{align}
	P_{\tilde{\delta}}(k) & = \abs{\Pi_{\zeta_{k}}}^2\left( \frac{1}{9\bar{a}^6 \Bar{H}^2(1+w)^2}\right),
\end{align}
where we used the flat Friedmann equation
\begin{align}
	\bar{H}^2 = \frac{\kappa \Bar{\rho}}{3}.
\end{align}
We are now able to compute the variance modes numerically using the above relation together with Eq.~\eqref{physicalrho} in the fluid gauge and our numerical code.

@@

\begin{figure}[tbp]
	\centering
	\includegraphics[width=.6\textwidth]{F.pdf}
	\caption{ Plot of the mass fraction of PBHs forming during bounce versus their formation mass in solar masses for $w = 10^{-10}$ and $x_b = 10^{35}$. The computations were done using the NumCosmo library.}
	\label{fvsm}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion and Conclusions}
\label{sec:discussion}
In this study, we investigated the formation of PBHs in a flat quantum bouncing model containing only dark matter with a small equation of state parameter $w$. Despite the spectra from this model exhibiting a slight blue tilt, which deviates from observations of the Cosmic Microwave Background (CMB), this discrepancy should not preclude the model, as the inclusion of radiation is expected to induce a different spectrum that may be compatible with observations \cite{Vitenti2012}. The dark matter-only model serves as an initial attempt to quantify PBH formation during the contracting phase.

@@

We may now compute $k$ using the second relation in Eq.~\eqref{kmdef}, which requires us to evaluate $\dot{R}(t, r)$. To do so, we define the inhomogenous Hubble function
\begin{align}
	\label{hdefltb}
	H & \equiv \frac{\dot{R}(t, r)}{R(t, r)}
	.\end{align}
Since we have the vanishing relations in Eqs.~\eqref{mdot} and \eqref{kdot}, $R$ is the
only local degree of freedom that evolves in time and provides the dynamics of the
universe according to Eq.~\eqref{ltb1}. Hence it is natural that we use this quantity to
define the Hubble function.  Assuming that the local metric is homogeneous at $t_{ini}$,
\begin{align}
	\label{hinicond}
	H^2(t_{ini}, r) & = \bar{H}^2(t_{ini}) = \frac{ \kappa \bar{\rho}}{3} = \frac{2 M}{1 +  \delta_{ini}} \frac{1}{r_c^3}
	.\end{align}
Thus, we can use this relation to compute $k$ for $t = t_{ini}$ since this variable does not evolve in time, which leads to
\begin{align}
	k(r) & = \frac{2m}{r} - r^2H^2_{ini}= \left\{\begin{array}{ll}
		                                             2M\frac{\delta_{ini}}{1+\delta_{ini}}\frac{r^2}{r_\mathrm{c}^3} & r \leq r_c \\
		                                             2M\frac{\delta_\mathrm{ini}}{1+\delta_{ini}}\frac{1}{r}         & r > r_c
	                                             \end{array}\right.
	.\end{align}

Now that we have properly computed our local quantities, we can rewrite our solutions in Eq.~\eqref{ltb1} and~\eqref{ltb2} for $r\leq r_c$ as
\begin{align}
	\label{ltb3}
	R(\theta, r)        & = \frac{r (1+\delta_{ini})}{\delta_{ini}}\sin^2\left(\frac{\theta}{2}\right),                             \\
	\label{ltb4}
	\Delta t(\theta,r ) & = \frac{1 + \delta_{ini}}{2H_{ini}\delta^{\frac{3}{2}}(t_{ini})}\left(\theta - \pi - \sin \theta \right),
\end{align}
such that $\Delta t = t_i - t_j$ for arbitrary times and $\theta_{ini}$ is obtained from Eq.~\eqref{ltb1} at $t_{ini}$, i.e,
\begin{align}
	\label{sintheta}
	\sin^2\left(\frac{\theta_{ini}}{2}\right) & = \frac{\delta_{ini}}{1 + \delta_{ini}}
	.\end{align}
We are now able to use these solutions to compute the critical value for the density contrast.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Fluid's Gauge}
\label{appgauge}
In the last appendix, we computed the solution for the Einstein equations for an inhomogeneous dust-perfect fluid with a spherically symmetric local metric. To use this solution and compare it with our perturbations defined in Sec.~\ref{linearpert}, we must ensure that measurements in the local metric are in the same Gauge as our perturbations. Since we projected the energy-momentum tensor in the direction of the velocity of the fluid in Eq.~\eqref{emprojec}, we consider the Gauge where the fluid is at rest and we shall now see how to define our perturbation theory in this Gauge choice.

@@

The first choice for $\mathcal{V}$ in Eq.~\eqref{psigauge} assures that the perturbed fluid is at rest with the background universe. The other two conditions guarantee that, at least for an initial time $t_1$, there are no off-diagonal terms and we have an isotropic perturbed fluid. For different times, it is not possible to set a Gauge such that $\mathcal{E}=0$ for $t \neq t_1$~\cite{vitenti2012large}. Nonetheless, we can argue that this term is proportional to the pressure of the fluid which has an almost vanishing value for cold dark matter and thus can be discarded. Also, since our main goal is to make proper measurements of the density contrast that depends on Eq.~\eqref{deltarhoinvariant}, this term is not relevant to our computations. Finally, in the fluid's Gauge, the gauge-invariant density contrast may be interpreted as its physical equivalent, that is,
\begin{align}
	\label{physicalrho}
	\Tilde{\delta \rho} & = \delta \rho + \mathcal{V}\dot{\Bar{\rho}} = \delta \rho\end{align}
where we used Eq.~\eqref{deltarhoinvariant}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\acknowledgments

>>>>

\section{Linear Scalar Perturbations}
\label{linearpert}

While the universe appears homogeneous and isotropic on large scales, the FLRW metric
falls short of providing a precise description of our universe, which presents
inhomogeneities that are associated with structure formation, e.g. galaxy clusters,
black holes, stars, and others. To characterize the physical universe, we study linear
scalar perturbations around our flat contracting background metric described in the
previous section. We will then investigate how said perturbations may seed the formation
of primordial black holes in the subsequent section. We will from now on characterize
background quantities with an overbar and perturbed (physical) quantities without it.

@@

Note from Eq.~\eqref{deltarhoinvariant} that if we choose a gauge where $\mathcal{V}
	=0$, $\Tilde{\delta\rho}$ becomes the physical density perturbation, which leads to an
easier interpretation of this quantity. Also, the choice of $\mathcal{V} =0$ will be
well suited to connect our perturbation theory with the Lemaitre-Toman-Bondi metric
discussed in App.~\ref{appgauge}.  Other works have defined the gauge invariant density
perturbation as
\begin{align}
	\label{newtoniandeltarho}
	\Tilde{\delta\rho}^{N} & \equiv \delta\rho - \delta\sigma \dot{\Bar{\rho}}
	,\end{align}
so that this quantity becomes the physical variable in the Newtonian Gauge, where
$\delta\sigma = 0$ and $\Phi = \phi$. However, in Ref.~\cite{vitenti2012large}, it was
shown that for bounce models with long contracting phases, regardless of the bounce
type, the Bardeen perturbation $\Phi$ grows larger than one and invalidates the
perturbative series, as $\phi$ also grows larger than one in the Newtonian Gauge.
Additionally, the definition in Eq.~\eqref{newtoniandeltarho} leads to
\begin{align}
	\zeta & = \Phi + \frac{2\bar{D}^2 \Phi}{3  \kappa (\bar{\rho}+\bar{p})} +
	\frac{{\tilde{\delta\rho}^N}} {3(\bar{\rho}+\bar{p})},
\end{align}
which implies that $\tilde{\delta\rho}^N$ grows with $\Phi$ and thus has larger spectra
as well as in this gauge. Hence, the Newtonian gauge is not a valid choice for bounce
models as it would lead to miss-calculation of the physical quantities whose values
would be inflated. We shall avoid this choice in this work and stick with the definition
in Eq.~\eqref{deltarhoinvariant}. We now compute the density contrast modes, which
require the perturbative equations of motion of the model.

@@

\subsection{Quantization}

@@

A widely used prescription to define such a vacuum state is known as the adiabatic
vacuum prescription~\cite{birrell1984quantum, mukhanov2007introduction}. Its main idea is
to set initial conditions by demanding that the mode functions $\zeta_{k}(t)$ coincide
with their ${\cal N}$ order adiabatic approximation, $^{({\cal N})}\zeta_{k}(t)$. This
is implemented if one chooses the initial conditions~\cite{birrell1984quantum}
\begin{align}
	\label{v_init}
	{ }^{(\mathcal{N})}	\zeta_k(t_0) & = \frac{1}{\sqrt{m(t_0)\omega_k}(t_0)}e^{i\alpha_k(t_0)}~\text{and}~ { }^{(\mathcal{N})}	\dot{\zeta}_k(t_0) = im(t_0)\omega_k(t_0)\zeta_k(t_0),
\end{align}
where $\omega_k^2(t) = \frac{c_s^2 k^2}{\bar{a}^2}$, $m(t) = 2z^2$ and $\alpha_k$ is
defined by the relation
\begin{align}
	{ }^{(\mathcal{N})}	\zeta_k(t_0) \dot{\alpha}_k(t_0) & = 1.
\end{align}

@@

As known from the literature, the dust bouncing model has an approximate scale-invariant
spectrum, that is
\begin{align}
	\label{powerlaw}
	P_{{\zeta}}(k) & \approx A k^{n_s-1}
	,\end{align}
where $A$ is a constant and $n_s$ is the spectral index. This is a general feature of
single barotropic fluid quantum bouncing models, and the spectral index is related to
the equation of state parameter $w$ by\footnote{For an explicit semi-analytic
	derivation, see \cite{nelson_peter_bouncing_original}.}
\begin{equation}
	n_{s}(w) = 1 + \frac{ 12w }{ 1 + 3w }\, ,
\end{equation}
%
which is nearly scale invariant for $|w| \ll 1 $. However, the usual positive values of
$w$ lead to a blue tilted spectrum $n_{s} > 1$, which differ from CMB observations
\cite{planck_inflation_constraints}, consistent with the inflationary scenario
prediction of a red tilted spectrum with $n_{s} \approx 0.96$. Although the initial
power spectrum is not consistent with observations, one must recall that our model is
considering a pure dark matter-dominated universe, neglecting the effects of radiation.
In particular, it has been suggested \cite{nelson2021bouncing} that the inclusion of
radiation may lead to a red almost scale-invariant spectrum. Therefore, this model must
be understood as a first approximation to a more complete model, which is still under
development. Furthermore, since most of the modes that influence our universe have
crossed the Hubble horizon during dust domination, this means that this model, even with
its simplicity, may still cover relevant information for the future complete model.

This result should not advocate for the prediction failure of the contracting scenario
since some works also suggest a blue-tilted spectrum for inflationary
models~\cite{Wang2014, Cai2015}. For instance, in Ref.~\cite{Kuroyanagi2021} they
discuss the possibility of explaining the recent NANOGrav results with a blue-tilted
spectra~\cite{Wu2023}. Also, the spectra are only dictated by a power law for narrow
intervals of momenta, which implies that corrections to Eq.~\eqref{powerlaw} must be
applied. Nonetheless, we see that the contracting phase produces a nearly
scale-invariant spectrum. In the next section, we begin to analyze the primordial black
hole formation seeded by the scalar perturbations in our cosmological background.

@@

It is indeed expected that a fluid with no pressure leads to a total collapse since no
forces are opposing the collapse and dust only interacts via gravity. Thus all dark
matter that we see today would indeed be contained in primordial black holes. However,
this hypothesis was already refuted by some works by constraining the PBH abundance in
dark matter based on observational effects (see Refs.~\cite{Villanueva2021, Carr2021}).
Hence we must consider a dust fluid with a non-vanishing equation of state.

\subsection{Filtered Variance}

Let us now compute the variance which will determine the distribution for PBH. Since
sub-Jeans and super-Hubble scales do not contribute to the PBH formation that we are
interested in, we want to compute the integral in  Eq.~\eqref{eqsigmar} between $k_H$
and $k_j$. To do so, we first need to evaluate $P_{\tilde{\delta}}(k)$ using
Eqs.~\eqref{deltarho}-\eqref{phiz} and the field expansions in Eqs.~\eqref{vexp}
and~\eqref{pvexp}. Explicitly,
\begin{align}
	\label{psiq}
	\hat{\Psi}(\eta,\textbf{x}) & =\int\frac{\mathrm{d}^3\textbf{k}}{(2\pi)^{\frac{3}{2}}}
	\frac{\bar{H}}{2\Bar{a}k^2}\left[\Pi_{\zeta_{k}} e^{i\textbf{k}\textbf{x}} a_{k} +
	\Pi_{\zeta_{k}}^{ *}e^{-i\textbf{k}\textbf{x}} a_{k}^{\dagger}\right]
	,\end{align}
\begin{align}
	\bar{D}^2\hat{\Psi}(\eta,\textbf{x}) & =
	-\int\frac{\mathrm{d}^3\textbf{k}}{(2\pi)^{\frac{3}{2}}}
	\frac{\bar{H}}{2a^3 }\left[\Pi_{\zeta_{k}} e^{i\textbf{k}\textbf{x}} a_{k} +
	\Pi_{\zeta_{k}}^{*}e^{-i\textbf{k}\textbf{x}} a_{k}^{\dagger}\right]
\end{align}
and
\begin{align}
	\hat{\tilde{\delta\rho}}(\textbf{x}) & =
	\int_{-\infty}^{\infty}\frac{\mathrm{~d}^3\textbf{k}}{(2\pi)^{\frac{3}{2}}}
	\left\{ \left(\frac{\bar{H}}{\kappa \bar{a}^3}\right)\Pi_{\zeta_{k}} e^{i\textbf{k}\textbf{x}} a_{k} +\right. \nonumber                                     \\
	                                     & \left. \left(\frac{\bar{H}}{\kappa a^3}\right)\Pi_{\zeta_{k}}^{*} e^{-i\textbf{k}\textbf{x}} a_{k}^{\dagger}\right\}
	.\end{align}
Thus, the two-point function and the variance are given by
\begin{align}
	\label{2point}
	\left<\hat{\Tilde{\delta}}(\textbf{x})\hat{\Tilde{\delta}}(\textbf{y})\right> & =
	\frac{1}{2\pi^2} \int_0^\infty \frac{\mathrm{~d}k}{k}~\left[P_{\tilde{\delta}}(k)
		\frac{\sin{kR}}{kR} \right]
\end{align}
and
\begin{align}
	P_{\tilde{\delta}}(k) & = \abs{\Pi_{\zeta_{k}}}^2\left( \frac{1}{9\bar{a}^6 \Bar{H}^2(1+w)^2}\right),
\end{align}
where we used the flat Friedmann equation
\begin{align}
	\bar{H}^2 = \frac{\kappa \Bar{\rho}}{3}.
\end{align}
We are now able to compute the variance modes numerically using the above relation
together with Eq.~\eqref{physicalrho} in the fluid gauge and our numerical code.

@@

\begin{figure}[tbp]
	\centering
	\includegraphics[width=.6\textwidth]{F.pdf}
	\caption{Plot of the mass fraction of PBHs forming during bounce versus their
		formation mass in solar masses for $w = 10^{-10}$ and $x_b = 10^{35}$. The
		computations were done using the NumCosmo library.}
	\label{fvsm}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion and Conclusions}
\label{sec:discussion}

In this study, we investigated the formation of PBHs in a flat quantum bouncing model
containing only dark matter with a small equation of state parameter $w$. Despite the
spectra from this model exhibiting a slight blue tilt, which deviates from observations
of the Cosmic Microwave Background (CMB), this discrepancy should not preclude the
model, as the inclusion of radiation is expected to induce a different spectrum that may
be compatible with observations \cite{Vitenti2012}. The dark matter-only model serves as
an initial attempt to quantify PBH formation during the contracting phase.

@@

We may now compute $k$ using the second relation in Eq.~\eqref{kmdef}, which requires us
to evaluate $\dot{R}(t, r)$. To do so, we define the inhomogenous Hubble function
\begin{align}
	\label{hdefltb}
	H & \equiv \frac{\dot{R}(t, r)}{R(t, r)}
	.\end{align}
Since we have the vanishing relations in Eqs.~\eqref{mdot} and \eqref{kdot}, $R$ is the
only local degree of freedom that evolves in time and provides the dynamics of the
universe according to Eq.~\eqref{ltb1}. Hence it is natural that we use this quantity to
define the Hubble function.  Assuming that the local metric is homogeneous at $t_{ini}$,
\begin{align}
	\label{hinicond}
	H^2(t_{ini}, r) & = \bar{H}^2(t_{ini}) = \frac{ \kappa \bar{\rho}}{3} = \frac{2 M}{1 +  \delta_{ini}} \frac{1}{r_c^3}
	.\end{align}
Thus, we can use this relation to compute $k$ for $t = t_{ini}$ since this variable does
not evolve in time, which leads to
\begin{align}
	k(r) & = \frac{2m}{r} - r^2H^2_{ini}= \left\{\begin{array}{ll}
		                                             2M\frac{\delta_{ini}}{1+\delta_{ini}}\frac{r^2}{r_\mathrm{c}^3} & r \leq r_c \\
		                                             2M\frac{\delta_\mathrm{ini}}{1+\delta_{ini}}\frac{1}{r}         & r > r_c
	                                             \end{array}\right.
\end{align}

Now that we have properly computed our local quantities, we can rewrite our solutions in
Eq.~\eqref{ltb1} and~\eqref{ltb2} for $r\leq r_c$ as
\begin{align}
	\label{ltb3}
	R(\theta, r)        & = \frac{r (1+\delta_{ini})}{\delta_{ini}}\sin^2\left(\frac{\theta}{2}\right),                             \\
	\label{ltb4}
	\Delta t(\theta,r ) & = \frac{1 + \delta_{ini}}{2H_{ini}\delta^{\frac{3}{2}}(t_{ini})}\left(\theta - \pi - \sin \theta \right),
\end{align}
such that $\Delta t = t_i - t_j$ for arbitrary times and $\theta_{ini}$ is obtained from
Eq.~\eqref{ltb1} at $t_{ini}$, i.e,
\begin{align}
	\label{sintheta}
	\sin^2\left(\frac{\theta_{ini}}{2}\right) & = \frac{\delta_{ini}}{1 + \delta_{ini}}
	.\end{align}
We are now able to use these solutions to compute the critical value for the density
contrast.

\section{Fluid's Gauge}
\label{appgauge}

In the last appendix, we computed the solution for the Einstein equations for an
inhomogeneous dust-perfect fluid with a spherically symmetric local metric. To use this
solution and compare it with our perturbations defined in Sec.~\ref{linearpert}, we must
ensure that measurements in the local metric are in the same Gauge as our perturbations.
Since we projected the energy-momentum tensor in the direction of the velocity of the
fluid in Eq.~\eqref{emprojec}, we consider the Gauge where the fluid is at rest and we
shall now see how to define our perturbation theory in this Gauge choice.

@@

The first choice for $\mathcal{V}$ in Eq.~\eqref{psigauge} assures that the perturbed
fluid is at rest with the background universe. The other two conditions guarantee that,
at least for an initial time $t_1$, there are no off-diagonal terms and we have an
isotropic perturbed fluid. For different times, it is not possible to set a Gauge such
that $\mathcal{E}=0$ for $t \neq t_1$~\cite{vitenti2012large}. Nonetheless, we can argue
that this term is proportional to the pressure of the fluid which has an almost
vanishing value for cold dark matter and thus can be discarded. Also, since our main
goal is to make proper measurements of the density contrast that depends on
Eq.~\eqref{deltarhoinvariant}, this term is not relevant to our computations. Finally,
in the fluid's Gauge, the gauge-invariant density contrast may be interpreted as its
physical equivalent, that is,
\begin{align}
	\label{physicalrho}
	\Tilde{\delta \rho} & = \delta \rho + \mathcal{V}\dot{\Bar{\rho}} = \delta \rho\end{align}
where we used Eq.~\eqref{deltarhoinvariant}.

\acknowledgments
